---
title: "SQL"
output: html_document
---

## Flitering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. Below you will find an example using the Lahman database. The first thing we'll need to do is load that database, as well as the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could write the following query: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

The next example you will see is a query looking for members of the New York Yankees that have hit up to 40 homeruns. For this, we would write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"
sqldf(query)
```


## ORDER BY

With SQL, you can order the rows of your output. For instance, suppose you want to see every instance of a player hitting more than 50 homeruns, but you would like the players with the most homeruns to be at the top. We could do the following:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

In the exmaple below, we are running a query searching for players that have struck out less than 10 times, have had at least 400 at-bats, and we are ordering the list with the least strikeouts at the top. For that, we would right the following:

```{r}
query<-"SELECT playerID,yearID,teamID,AB,SO FROM Batting
WHERE SO<10 and AB>=400

ORDER BY SO"

sqldf(query)
```

## Aggregation

Sometimes when working with SQL, there will be a need to condense results to the minimum possible output. Below you will find a couple examples of how we can do this, with my examples using sum(HR) and avg(HR).

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)
```

As you can see above, we queried any players that have 600 or more homeruns for their whole careers, putting them in order from greatest to least. In the second example, we were looking for players who in their careers averaged more than 30 homeruns, also in descending order.

## Joins

Sometimes when working with SQL, there are instances where you want to join multiple tables to come together as one output. Below I will post a couple examples of code used to go about this.

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

In a future example, we will go over how to join three tables together.

## JOINS UPDATE September 13, 2017

As stated in the previous update of the website, I now have an instance of code where you would be able to link three tables together. The example is as follows:

```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

As you can see above, there are two INNER JOIN commands in the syntax of our code, one of which links the Teams table and the other linking the Master table, making connections with teamID, yearID, and playerID.