---
title: "SQL"
output: html_document
---

## Flitering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. Let's look at an example using the Lahman database. So we should first load that database along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could write the following query: 

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

```{r}
query<-"SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"
sqldf(query)
```


## ORDER BY

With SQL, you can order the rows of your output. For instance, suppose you want to see every instance of a player hitting more than 50 homeruns, but you would like the players with the most homeruns to be at the top. We could do the following:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

